
\begin{thebibliography}{99}



\bibitem{AfD} Miguel Cepero, (2014), \emph{Appetite for Destruction},\\ Available: http://procworld.blogspot.co.uk/2014/08/appetite-for-destruction.html,\\ Last accessed 22th Oct 2014.

\bibitem{OCL} Miguel Cepero, (2011), \emph{OpenCL Voxelization},\\ Available: http://procworld.blogspot.co.uk/2011/04/opencl-voxelization.html,\\ Last accessed 22th Oct 2014.

\bibitem{PW} Miguel Cepero, (2014), \emph{Procedural World},\\ Available: http://procworld.blogspot.co.uk/,\\ Last accessed 22th Oct 2014.

\bibitem{MCGPU} Scrawk, (2014), \emph{Marching Cubes on the GPU in Unity},\\ Available: http://scrawkblog.com/2014/10/16/marching-cubes-on-the-gpu-in-unity/,\\ Last accessed 22th Oct 2014.

\bibitem{SW} \emph{Fast Parallel Surface and Solid Voxelization on GPUs},\\ M. Schwarz, H. Seidel, \emph{ACM Transactions on Graphics}, Dec 2010, pp.~179:1--179:9.

\bibitem{V} \emph{A Low Cost Antialiased Space Filled Voxelization Of Polygonal Objects
},\\ S. Thon, G. Gesquière, R. Raffin, \emph{Graphicon}, 2004.

\bibitem{MC} \emph{Marching Cubes: A High Resolution 3D Surface Construction Algorithm
},\\ W. Lorensen, H. Cline, \emph{SIGGRAPH Comput. Graph.}, July 1987, pp.~163--169.

\bibitem{RTC} \emph{Real-time Collision Detection for Dynamic Hardware Tessellated Objects
},\\ M. Nie{\ss}ner et al, \emph{Eurographics}, 2013.

\bibitem{RTCV} M. Nie{\ss}ner et al, (2013), \emph{Real-time Collision Detection for Dynamic Hardware Tessellated Objects}, [Online video],\\ Available: \href{http://graphics.stanford.edu/~niessner/papers/2013/2collision/niessner2013collision.mp4}{http://graphics.stanford.edu/$\sim$niessner/papers/2013/}\\\href{http://graphics.stanford.edu/~niessner/papers/2013/2collision/niessner2013collision.mp4}{2collision/niessner2013collision.mp4},\\ Last accessed 23th Oct 2014.

\end{thebibliography}
