
\section{Timeline}

\subsection*{Sprint 1 (24/10/14--7/11/14)}

Implement a mesh to voxel transformation algorithm from section \ref{methods} and become familiar with the voxel to mesh transformation plugin.

Milestones: Some example code, which will probably not be
used in the final project, and some example scenes demonstrating conversion.


\subsection*{Sprint 2 (7/11/14--21/11/14)}

Refactor the mesh to voxel algorithm to only voxelise a portion of an object.

Milestones: Working voxelisation of predefined portions of an object.


\subsection*{Sprint 3 (28/11/14--5/12/14)}

Hook the voxelisation algorithm to collision cues above a predefined threshold.

Milestone: Examples of conversion occurring and not occurring as required on cue.


\subsection*{Sprint 4 (5/12/14--16/1/15)}

Review code and progress over the Christmas break. Ensure everything is well documented, including source and evaluation of current progress.


\subsection*{Sprint 5 (16/1/15--30/1/15)}

Write the initial chapters of the dissertation. Write the progress report. Build the progress demonstration as detailed in section \ref{demo}

Milestones: Preparation chapter of Dissertation complete. Submission of progress report. Demonstration ready.


\subsection*{Sprint 6 (30/1/15--13/2/15)}

Begin to work on physically based destruction algorithms for the splitting apart of volumetric objects, including transferral of force to any objects formed from dissociated voxels. Define the parameters which will impact physical reactions.

Milestones: Ability to define physical properties for game objects.

\clearpage
\subsection*{Sprint 7 (13/2/15--27/2/15)}

Complete work on physically based destruction algorithms.

Milestones: Data confirming impact propagation through voxel bonds.

\subsection*{Sprint 8 (27/2/15--13/3/15)}

Integrate the voxel to mesh plugin with the result of volumetric destruction.

Milestones: Able to see object split apart as the result of an impact of the correct magnitude.


\subsection*{Sprint 9 (13/3/15--27/3/15)}

Implement real-time user modification of objects.

Milestones: Example scene where playing is able to modify objects via physical and non-physical means.


\subsection*{Sprint 10 (27/3/15--10/4/15)}

Code review and code freeze. All success criteria should be implemented by this point. No new features will be written during this period, only bug fixes and optimisations.\\
Further dissertation progress will be made.


\subsection*{Sprint 11 (10/4/15--24/4/15)}

Create comparable scenes using base Unity3D physics for profiling.\\
Profile against these scenes as well as on computers with varying specifications.


\subsection*{Sprint 12 (24/4/15--8/5/15)}

Finish the dissertation, preparing graphs for insertion.  Review the whole
project, check the dissertation, and spend a final few days on
whatever is in greatest need of attention.


\subsection*{Final Sprint (8/5/15--15/5/15)}

Final proofreading and submission of dissertation.

Milestone: Submission of Dissertation. 
